### 2020-02-12 paper4/article.tex
DLCD -> HLGD. Re-read Chp 2,3. Briefly paper 7.

We demonstrate here that when a distant wall is considered, the evolution of $\bm r$ remains autonomous in the leading approximation, $\dot{\bm r}=\bm V(\bm r(t))$. Thus at any finite $z_0\gg a$  we can still examine the phase portrait, which is however qualitatively different from that in Fig.~\ref{fig:SeparatrixBG}. Our calculation is not a perturbation theory of the BG solution as we do not assume $\bm V\approx \bm V^0$, so the disturbed phase space flow $\bm V(\bm r)$ is significantly different from $\bm V^0$. The change in topology occurs because at finite $z_0$ there exist critical points at which $\bm V(\bm r)$ vanishes, see the phase portrait in the symmetry plane in Fig.~\ref{fig:Schematics}. The saddle (hyperbolic) point $r_s$, the closer of the two critical points to the origin, is unstable. The other neutral equilibrium (elliptic) point $r_c$ corresponds to a marginally stable configuration, where the pair flows without changing its inter-particle distance and orientation, see Fig.~\ref{Draw1}(a). Not too large deviations from this state result in the sphere orbiting around this elliptic point. These dancing trajectories would have rather unusual appearance when considered in the laboratory frame: while one sphere travels downstream, the other sphere revolves around a point co-moving in space with the first sphere, see Fig.~\ref{Draw1}(b). The phase plane at $x>0$ is characterized by two disconnected regions of closed trajectories shown in Fig.~\ref{fig:Schematics} by the blue curves. The region to the left of
the separatrix (blue) that crosses $x=r_s$ resembles the BG's closed trajectories. The trajectories around the elliptic point $(r_c, 0)$ are solely due to the presence of the wall. The lowest red curve is an open trajectory similar to the BG's: after the encounter the vertical separation is restored to its initial value. In contrast to that, along the (red) trajectories circumventing the elliptic point, the vertical separation of the particles reverses sign after the encounter. We call these ``swapping trajectories", since they seem to correspond to the numerical findings of \cite{zurita-gotor_2007} at $z_0\sim a$, where trajectories with particles swapping their vertical positions after the encounter in a channel flow were reported (careful consideration of the Figure presented in \cite{zurita-gotor_2007} reveals slight changes of the vertical coordinates which seem to be a higher order effect than that considered here). The phase portrait for the evolution of inter-particle distance along the swapping trajectory obtained in \cite{zurita-gotor_2007} numerically, agrees with that predicted here theoretically. To prove that the sign-reversal of the vertical separation predicted here implies swapping, it has to be shown that the center of mass of the pair is not displaced vertically as a result of the encounter. We leave the rigorous proof for future work, focusing here on the evolution of inter-particle distance only. Thus, the use of the term ``swapping trajectories" here, strictly speaking, refers to open trajectories with sign-reversal of the vertical separation following the encounter.

@@

Here we derive the integral equation that determines the particle velocities. For future generalization to the case where the distances from the spheres to both walls are comparable we perform the derivation starting from the full formulation given by Eq.~(\ref{ds}).

@@

\begin{figure}
\begin{tabular}{ccc}
\includegraphics[width=0.43\textwidth]{Fig4a.pdf} \hspace{1cm} &
\includegraphics[width=0.43\textwidth]{Fig4b.pdf}
\\
(a) & (b)
\end{tabular}
\caption{The phase portrait in the $xz$-plane for $z_0=5$ (a) and,
$z_0=10$ (b). The BG separatrix obeying $z^2=16/(9x^3)$ at large $x$ is depicted by the dashed (grey) line. For any finite $z_0$, the phase portrait contains two disconnected regions of closed trajectories, in contrast to one region at $z_0=\infty$. Region I, where all trajectories are closed and the spheres are close to each other, is similar to that at $z_0=\infty$. Region II is also similar to the $z_0=\infty$ case: all the trajectories are open and the vertical separation after the interaction returns to its original value.
Region III has no counterpart at $z_0=\infty$. This region contains both closed and open trajectories (see Fig.~\ref{Separatrix2D} for a more detailed description). The trajectories passing not far from the stationary point $r_c$ are closed, orbiting around this point. The swapping open trajectories instead are characterized by a sign reversal of the vertical component of the separation vector after the encounter. The region of swapping trajectories is bounded from one
side by the closed trajectories around $r_c$ and from the other side by open non-swapping trajectories.}
\label{fig:Fig4a}
\end{figure}

To construct the separatrices (defined here as curves separating regions of qualitatively different behavior) in the $xz$-plane for given value of $z_0$ we first
find the critical point on the $x$-axis $(r_s,0,0)$ where the approximate value of $r_s$ is given in Section III.
One separatrix (red curve in Fig. \ref{fig:Fig4a}) is stable, see Fig.~\ref{fig:Schematics} and thus is computed
using integration of the original equations. The other separatrix (blue curve in Fig. \ref{fig:Fig4a}) is unstable as seen from Fig. \ref{fig:Schematics}. Thus, it is
found by
backward integration in time, for which it is stable, until the trajectory reaches the $z$-axis. All the trajectories below the blue curve (region I) are closed, while
those between the red and blue curves (region II) are open -- they correspond to
non-swapping trajectories. The trajectories between the red curve and the $x$-axis (region III)
can be divided into two classes -- open swapping trajectories (brown, black curves in the inset of Fig.~\ref{Separatrix2D}) and closed trajectories characterized by a
very large separation between the spheres (green, blue curves), see the captions of the Figures and detailed theory in the next Section.

@@

First we observe that the evolution of trajectories in the dancing-swapping region III admits $r\gg 1$ and thus can be simplified. It is readily seen numerically that, at least for $z_0\geq 5$ that are of interest here, we have $V^0_x(\bm r)\approx z$ within less than 15\% accuracy, meaning that the BG velocity difference is fully determined by the undisturbed shear flow. This is because the hydrodynamic interactions' correction to $V^0_x(\bm r)$ decays quickly with the spheres' separation, see Eqs.~(\ref{practical}), (\ref{largr}). We find from Eq.~(\ref{practical}) that we can use $V_x(\bm r)\approx z$ everywhere in region III. Moreover, we observe that $\delta V_z(x, 0, z)-\delta V_z(x, 0, 0)$ grows quadratically with $z$, see Appendix \ref{cartesian}. It is then found that since small-$z$ approximation holds (see below) then we can then use $\delta V_z(x, 0, z)\approx \delta V_z(x, 0, 0)$ in the whole region III. Finally, we can use the reduced Eq.~(\ref{velsc}). We find that the evolution of the trajectories in region III can be accurately described by the reduced system of equations,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
\dot x=z,\ \ \dot z=-\frac{8x }{3r^5}\!-\!\frac{5 xz^2}{r^5}+\frac{10x z_0^2(16z_0^2\!-\!x^2)}{(x^2+4z_0^2)^{7/2}},
\label{ds1}
\end{eqnarray}
where we assumed $z_0\geq 5$ and used Eq.~(\ref{bgz}). Furthermore, since region III is characterized by small $z$ then it is seen that $r\approx x$ and the second term in the RHS of the equation on $z$ can be dropped. Indeed, the ratio $15 z^2/8$ of this term to the first term in the RHS is small at moderate $x$ and not so small at larger $x$. However, at larger $x$ the time-derivative $\dot z$ is determined by the third term. Thus the second term is uniformly small everywhere in III as we verified numerically, and Eq.~(\ref{ds1}) is rewritten as
\begin{equation}
\dot x = z,
\quad
\dot z=-\frac{8}{3x^4}+
\frac{10x z_0^2(16z_0^2-x^2)}{(x^2+4z_0^2)^{7/2}}.
\label{ds2}
\end{equation}
The trajectories produced by this system in region III are indistinguishable from those produced by the full $\bm V(\bm r)$. The critical points of this approximate evolution obviously coincide with those obtained in Eq.~(\ref{critical_pts_eq}), so that, e. g. $10r_s z_0^2(16z_0^2-r_s^2)/(r_s^2+4z_0^2)^{7/2}=8/(3r_s^4)$ with $r_s= \left(32 z_0^3/15\right)^{1/5}$. Eliminating time variable we arrive at
\begin{equation}
\frac{d}{dx}z^2=-\frac{16}{3x^4}+
\frac{20x z_0^2(16z_0^2-x^2)}{(x^2+4z_0^2)^{7/2}}.
\label{ds3}
\end{equation}
The trajectory that crosses the $x$-axis at $x=x_i$ is given by the solution of the above equation and it reads
\begin{equation}
z^2=
\frac{16}{27x^3}
+\frac{20z_0^2(x^2-8z_0^2)}{3(x^2+4z_0^2)^{5/2}}
-\frac{16}{27x_i^3}
-\frac{20z_0^2(x_i^2-8z_0^2)}{3(x_i^2+4z_0^2)^{5/2}}.
\label{ds4}
\end{equation}
Setting here $x_i=r_s$ and using the condition on $r_s$ provided after Eq.~(\ref{ds2}), we arrive at the equation of the separatrix bounding the dancing-swapping region and separating it from region II (red line in Fig. \ref{fig:Fig4a})
\begin{equation}
z_{II, III}^2=
\frac{16}{27 x^3}
+\frac{20z_0^2(x^2-8z_0^2)}{3(x^2+4z_0^2)^{5/2}}+\frac{40z_0^2\left(48z_0^4-2r_s^2 z_0^2-r_s^4\right)}{9(r_s^2+4z_0^2)^{7/2}},\ \ r_s= \left(\frac{32 z_0^3}{15}\right)^{1/5}.
\label{ds5}
\end{equation}
We further find the following asymptotic behavior
\begin{equation}
z_{II, III}^2(x=\infty) =\frac{40z_0^2\left(48z_0^4-2r_s^2 z_0^2-r_s^4\right)}{9(r_s^2+4z_0^2)^{7/2}}\approx \frac{5}{3z_0},
\end{equation}
where the first equality holds down to $z_0=5$ and the last equality assumes $z_0\gg 1$. The equations confirm that region III has a finite width in $z$-direction, the fact underlying the validity of $\delta V_z(x, 0, z)\approx \delta V_z(x, 0, 0)$. The last equality provides the scaling law of growth of region III as the proximity to the wall decreases from $z_0=\infty$ to some finite value.

>>>>

We demonstrate here that when a distant wall is considered, the evolution of $\bm r$ remains autonomous in the leading approximation, $\dot{\bm r}=\bm V(\bm r(t))$. Thus at any finite $z_0\gg a$  we can still examine the phase portrait, which is however qualitatively different from that in Fig.~\ref{fig:SeparatrixBG}. Our calculation is not a perturbation theory of the BG solution as we do not assume $\bm V\approx \bm V^0$, so the disturbed phase space flow $\bm V(\bm r)$ is significantly different from $\bm V^0$. The change in topology occurs because at finite $z_0$ there exist critical points at which $\bm V(\bm r)$ vanishes, see the phase portrait in the symmetry plane in Fig.~\ref{fig:Schematics}. The saddle (hyperbolic) point $r_s$, the closer of the two critical points to the origin, is unstable. The other neutral equilibrium (elliptic) point $r_c$ corresponds to a marginally stable configuration, where the pair flows without changing its inter-particle distance and orientation, see Fig.~\ref{Draw1}(a). Not too large deviations from this state result in the sphere orbiting around this elliptic point. These dancing trajectories would have rather unusual appearance when considered in the laboratory frame: while one sphere travels downstream, the other sphere revolves around a point co-moving in space with the first sphere, see Fig.~\ref{Draw1}(b). The phase plane at $x>0$ is characterised by two disconnected regions of closed trajectories shown in Fig.~\ref{fig:Schematics} by the blue curves. The region to the left of
the separatrix (blue) that crosses $x=r_s$ resembles the BG's closed trajectories. The trajectories around the elliptic point $(r_c, 0)$ are solely due to the presence of the wall. The lowest red curve is an open trajectory similar to the BG's: after the encounter the vertical separation is restored to its initial value. In contrast to that, along the (red) trajectories circumventing the elliptic point, the vertical separation of the particles reverses sign after the encounter. We call these ``swapping trajectories", since they seem to correspond to the numerical findings of \cite{zurita-gotor_2007} at $z_0\sim a$, where trajectories with particles swapping their vertical positions after the encounter in a channel flow were reported (careful consideration of the Figure presented in \cite{zurita-gotor_2007} reveals slight changes of the vertical coordinates which seem to be a higher order effect than that considered here). The phase portrait for the evolution of inter-particle distance along the swapping trajectory obtained in \cite{zurita-gotor_2007} numerically, agrees with that predicted here theoretically. To prove that the sign-reversal of the vertical separation predicted here implies swapping, it has to be shown that the center of mass of the pair is not displaced vertically as a result of the encounter. We leave the rigorous proof for future work, focusing here on the evolution of inter-particle distance only. Thus, the use of the term ``swapping trajectories" here, strictly speaking, refers to open trajectories with sign-reversal of the vertical separation following the encounter.

@@

Here we derive the integral equation that determines the particle velocities. For future generalisation to the case where the distances from the spheres to both walls are comparable we perform the derivation starting from the full formulation given by Eq.~(\ref{ds}).

@@

\begin{figure}
\begin{tabular}{ccc}
\includegraphics[width=0.43\textwidth]{Fig4a.pdf} \hspace{1cm} &
\includegraphics[width=0.43\textwidth]{Fig4b.pdf}
\\
(a) & (b)
\end{tabular}
\caption{The phase portrait in the $xz$-plane for $z_0=5$ (a) and,
$z_0=10$ (b). The BG separatrix obeying $z^2=16/(9x^3)$ at large $x$ is depicted by the dashed (grey) line. For any finite $z_0$, the phase portrait contains two disconnected regions of closed trajectories, in contrast to one region at $z_0=\infty$. Region I, where all trajectories are closed and the spheres are close to each other, is similar to that at $z_0=\infty$. Region II is also similar to the $z_0=\infty$ case: all the trajectories are open and the vertical separation after the interaction returns to its original value.
Region III has no counterpart at $z_0=\infty$. This region contains both closed and open trajectories (see Fig.~\ref{Separatrix2D} for a more detailed description). The trajectories passing not far from the stationary point $r_c$ are closed, orbiting around this point. The swapping open trajectories instead are characterised by a sign reversal of the vertical component of the separation vector after the encounter. The region of swapping trajectories is bounded from one
side by the closed trajectories around $r_c$ and from the other side by open non-swapping trajectories.}
\label{fig:Fig4a}
\end{figure}

To construct the separatrices (defined here as curves separating regions of qualitatively different behavior) in the $xz$-plane for given value of $z_0$ we first
find the critical point on the $x$-axis $(r_s,0,0)$ where the approximate value of $r_s$ is given in Section III.
One separatrix (red curve in Fig. \ref{fig:Fig4a}) is stable, see Fig.~\ref{fig:Schematics} and thus is computed
using integration of the original equations. The other separatrix (blue curve in Fig. \ref{fig:Fig4a}) is unstable as seen from Fig. \ref{fig:Schematics}. Thus, it is
found by
backward integration in time, for which it is stable, until the trajectory reaches the $z$-axis. All the trajectories below the blue curve (region I) are closed, while
those between the red and blue curves (region II) are open -- they correspond to
non-swapping trajectories. The trajectories between the red curve and the $x$-axis (region III)
can be divided into two classes -- open swapping trajectories (brown, black curves in the inset of Fig.~\ref{Separatrix2D}) and closed trajectories characterised by a
very large separation between the spheres (green, blue curves), see the captions of the Figures and detailed theory in the next Section.

@@

First we observe that the evolution of trajectories in the dancing-swapping region III admits $r\gg 1$ and thus can be simplified. It is readily seen numerically that, at least for $z_0\geq 5$ that are of interest here, we have $V^0_x(\bm r)\approx z$ within less than 15\% accuracy, meaning that the BG velocity difference is fully determined by the undisturbed shear flow. This is because the hydrodynamic interactions' correction to $V^0_x(\bm r)$ decays quickly with the spheres' separation, see Eqs.~(\ref{practical}), (\ref{largr}). We find from Eq.~(\ref{practical}) that we can use $V_x(\bm r)\approx z$ everywhere in region III. Moreover, we observe that $\delta V_z(x, 0, z)-\delta V_z(x, 0, 0)$ grows quadratically with $z$, see Appendix \ref{cartesian}. It is then found that since small-$z$ approximation holds (see below) then we can then use $\delta V_z(x, 0, z)\approx \delta V_z(x, 0, 0)$ in the whole region III. Finally, we can use the reduced Eq.~(\ref{velsc}). We find that the evolution of the trajectories in region III can be accurately described by the reduced system of equations,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
\dot x=z,\ \ \dot z=-\frac{8x }{3r^5}\!-\!\frac{5 xz^2}{r^5}+\frac{10x z_0^2(16z_0^2\!-\!x^2)}{(x^2+4z_0^2)^{7/2}},
\label{ds1}
\end{eqnarray}
where we assumed $z_0\geq 5$ and used Eq.~(\ref{bgz}). Furthermore, since region III is characterised by small $z$ then it is seen that $r\approx x$ and the second term in the RHS of the equation on $z$ can be dropped. Indeed, the ratio $15 z^2/8$ of this term to the first term in the RHS is small at moderate $x$ and not so small at larger $x$. However, at larger $x$ the time-derivative $\dot z$ is determined by the third term. Thus the second term is uniformly small everywhere in III as we verified numerically, and Eq.~(\ref{ds1}) is rewritten as
\begin{equation}
\dot x = z,
\quad
\dot z=-\frac{8}{3x^4}+
\frac{10x z_0^2(16z_0^2-x^2)}{(x^2+4z_0^2)^{7/2}}.
\label{ds2}
\end{equation}
The trajectories produced by this system in region III are indistinguishable from those produced by the full $\bm V(\bm r)$. The critical points of this approximate evolution obviously coincide with those obtained in Eq.~(\ref{critical_pts_eq}), so that, e. g. $10r_s z_0^2(16z_0^2-r_s^2)/(r_s^2+4z_0^2)^{7/2}=8/(3r_s^4)$ with $r_s= \left(32 z_0^3/15\right)^{1/5}$. Eliminating time variable we arrive at
\begin{equation}
\frac{d}{dx}z^2=-\frac{16}{3x^4}+
\frac{20x z_0^2(16z_0^2-x^2)}{(x^2+4z_0^2)^{7/2}}.
\label{ds3}
\end{equation}
The trajectory that crosses the $x$-axis at $x=x_i$ is given by the solution of the above equation and it reads
\begin{equation}
z^2=
\frac{16}{27x^3}
+\frac{20z_0^2(x^2-8z_0^2)}{3(x^2+4z_0^2)^{5/2}}
-\frac{16}{27x_i^3}
-\frac{20z_0^2(x_i^2-8z_0^2)}{3(x_i^2+4z_0^2)^{5/2}}.
\label{ds4}
\end{equation}
Setting here $x_i=r_s$ and using the condition on $r_s$ provided after Eq.~(\ref{ds2}), we arrive at the equation of the separatrix bounding the dancing-swapping region and separating it from region II (red line in Fig. \ref{fig:Fig4a})
\begin{equation}
z_{II, III}^2=
\frac{16}{27 x^3}
+\frac{20z_0^2(x^2-8z_0^2)}{3(x^2+4z_0^2)^{5/2}}+\frac{40z_0^2\left(48z_0^4-2r_s^2 z_0^2-r_s^4\right)}{9(r_s^2+4z_0^2)^{7/2}},\ \ r_s= \left(\frac{32 z_0^3}{15}\right)^{1/5}.
\label{ds5}
\end{equation}
We further find the following asymptotic behavior
\begin{equation}
z_{II, III}^2(x=\infty) =\frac{40z_0^2\left(48z_0^4-2r_s^2 z_0^2-r_s^4\right)}{9(r_s^2+4z_0^2)^{7/2}}\approx \frac{5}{3z_0},
\end{equation}
where the first equality holds down to $z_0=5$ and the last equality assumes $z_0\gg 1$. The equations confirm that region III has a finite width in $z$-direction, the fact underlying the validity of $\delta V_z(x, 0, z)\approx \delta V_z(x, 0, 0)$. The last equality provides the scaling law of growth of region III as the proximity to the wall decreases from $z_0=\infty$ to some finite value.
